Generative AI for image reconstruction\\ in Intensity Interferometry: a first attempt

### 2025-09-24 Phase_II/draft/response.txt

|  - gravitational darkening -> gravity darkening

Modified

|  - The sentence referring to the "ground truth, which the Discriminator uses
|  to distinguish between real images and those generated by the Generator.
|  During training, the GAN aims to replicate these ground truth images."
|  seems to be confused about GANs in theory. The references to this
|  throughout the paper, including in the context of the merged data, need to
|  be clearer. The discriminator doesn't have access to the ground truth, the
|  discriminator has to distinguish between simulated and ground-truth data.

It has been rephrased and clear now. Yes, the Discriminator doesn't have access to the ground truth.
Even Discriminator loss function use Ground truth and generated images to give feedback to the Discriminator

@@

|  - There appears to be no correlation between the x and y centroids of the
|  star relative to the ground truth. This would be a serious problem? It
|  seems to do ok on the second moments, but no recovery at all of the third
|  moments. This may be related to section 5.3 which seems to lack any account
|  of gravity darkening, which is in the simulations.

Counter-argument: yes the centroids are showing noise -- all values
are around (63.5,63.5) which is the centre of the field -- but it is
not important, because absolute centroid is not a measurable quantity
in interferometry.

|  - No information is provided in the paper about computer systems. What
|  hardware was used (CPUs? GPUs?), and how long did it take for training and
|  for inference? It is important to quantify computational performance.

The information is provided and we use CPU for training. This work has used 2 nodes with 96 threads each,
which takes around 9 hour to train the model.

### 2025-09-24 Phase_II/draft/ParaGan.tex
first small modification

\section{Network Parameters}
\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
\caption{Example results of image reconstruction using the GAN model along with the II observations simulated in this work. Each row in this figure represents the results for a hypothetical fast-rotating star. Going from left to right in each row, the first panel represents the simulated $(u,v)$ plane II signals obtained using the six baselines (of Fig.\ref{fig:ft_base}). This image is presented, as the input, to the trained GAN to produce a predicted image. The second panel is the real image of the star, also called the ground truth. The third panel is the reconstructed image, or the predicted image, produced by the trained GAN model . The fourth and the last panel is the difference between the ground truth and the predicted image in the $(u,v)$ plane. The uniform background color in the difference panel suggests that the network is picking up white noise from the $(u,v)$ plane. }
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters of the GAN architecture used for reconstructing images of stellar objects using II. Given the adversarial nature of GANs—where the Generator and Discriminator engage in a min-max game—careful tuning of key parameters is critical to ensure that both networks are well-balanced for effective training.

>>>>

\section{Network Parameters}
\begin{figure*}
	\centering
	\includegraphics[width=\linewidth]{fig/testing_image/image_1.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_3.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_4.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_10.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_12.png}
	\includegraphics[width=\linewidth]{fig/testing_image/image_15.png}
\caption{Example results of image reconstruction using the GAN model along with the II observations simulated in this work. Each row in this figure represents the results for a hypothetical fast-rotating star. Going from left to right in each row, the first panel represents the simulated $(u,v)$ plane II signals obtained using the six baselines (of Fig.\ref{fig:ft_base}). This image is presented, as the input, to the trained GAN to produce a predicted image. The second panel is the real image of the star, also called the ground truth. The third panel is the reconstructed image, or the predicted image, produced by the trained GAN model . The fourth and the last panel is the difference between the ground truth and the predicted image in the $(u,v)$ plane. The uniform background color in the difference panel suggests that the network is picking up white noise from the $(u,v)$ plane. }
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters of the GAN architecture used for reconstructing images of stellar objects using II. Given the adversarial nature of GANs—where the Generator and Discriminator engage in a min-max game—careful tuning of key parameters is critical to ensure that both networks are well-balanced for effective training.

### 2025-09-24 Phase_II/draft/Result.tex
first small modification

\section{Images Reconstructed by the GAN}
In this section, we begin by discussing phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis in which multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of \(2 \cdot 10^{-4}\), a kernel size of 5×5, and equal noise percentages applied to both the original and generated images. A batch size of 1 is used, and equal training is provided to both the Discriminator and the Generator.

\subsection{Predicted Image from the Trained GAN}
Fig.~\ref{fig:GAN} demonstrates the success of the GAN in training a model for Intensity Interferometry (II) to reconstruct images of fast-rotating stars. The GAN was trained on the training datasets for 60,000 steps and subsequently tested on various validation datasets to produce predicted images of a fast-rotating star. In Fig.~\ref{fig:GAN}, four combined images illustrate the GAN's performance in reconstructing the star’s shape, size, and brightness distribution using II.
\begin{itemize}
\item{The left panel shows the signals collected from six baselines, which serve as the input for the Generator during training.}
\item{The first middle panel displays the real image, or ground truth, which the Discriminator uses to distinguish between real images and those generated by the Generator. During training, the GAN aims to replicate these ground truth images.}
\item{The second middle panel presents the reconstructed, or predicted, image produced by the trained GAN, highlighting its success in image reconstruction.}
\item{The right panel shows the difference between the ground truth and the predicted image, with a smaller difference indicating higher precision in image reconstruction.}
\end{itemize}
The predicted images in Fig.~\ref{fig:GAN} yield encouraging results, accurately conveying visual information about the source's size, shape, and brightness distribution across its surface using only six baselines. However, further improvements can be achieved by increasing the number of telescopes to maximize coverage of the $(u, v)$ plane, making the existing and upcoming CTAO an ideal candidate for this approach.

@@

The raw moment $M_{ij}$ of an image $I(x, y)$ is defined as \citep{hu1962visual}
\begin{equation}
	M_{ij} = \sum_{x} \sum_{y} x^i y^j I(x, y).
	\label{eqn:Mom}
\end{equation}
The zeroth order raw moment, or monopole, represents the total intensity of an image. It is computed by summing all pixel values across the image, yielding an overall intensity measure. In this context, analyzing the monopole provides the total flux of fast-rotating stars. According to Eq.~\eqref{eqn:Mom}, the monopole of an image is calculated as:
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
The left figure of Fig.~\ref{fig:cen} displays the monopole values for 50 reconstructed images. The plot reveals a linear relationship between the monopole of the ground truth (real image) on the $x$-axis and that of the predicted (reconstructed) image on the $y$-axis, consistent across sources of varying shapes and sizes. This linearity confirms that the predicted images have an overall intensity (flux) that closely matches the ground truth. However, while the monopole effectively represents the total brightness, it does not provide information about the position, shape, size, or detailed brightness distribution of the fast-rotating stars. For these aspects, higher-order moments are necessary.

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid ($x$-centroid and $y$-centroid). It represents the spatial position of the image and is calculated using first-order raw moment and monopole. The formulation of centroid along the $x$ and $y$ directions is
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
The middle and right figure of Fig.~\ref{fig:cen} show the comparison of the $x$-centroid and $y$-centroid for 50 predicted images with respect to ground truths, respectively. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star compared with the ground truth.

The center of mass of a fast-rotating star, or any stellar object, is determined by its centroid, which provides the $x$ and $y$ coordinates representing the spatial position of the image. This centroid is computed using the first-order raw moments in conjunction with the monopole (the zeroth-order moment). The formulations for the centroid along the $x$ and $y$ directions are given by:
\begin{eqnarray}
&&x_c = \frac{M_{10}}{M_{00}} = \frac{\sum_{x,y} x \cdot I(x,y)}{\sum_{x,y} I(x,y)} \nonumber \\
&&y_c = \frac{M_{01}}{M_{00}} = \frac{\sum_{x,y} y \cdot I(x,y)}{\sum_{x,y} I(x,y)}
\end{eqnarray}
Here, \(I(x,y)\) represents the intensity at pixel \((x,y)\), \(M_{00}\) is the monopole (total intensity), and \(M_{10}\) and \(M_{01}\) are the first-order raw moments along the $x$ and $y$ axes, respectively. This formulation accurately captures the spatial center of mass of the stellar object in the image.

The middle and right panel of Fig.~\ref{fig:cen} compare the centroids ($x_{c}$, $y_{c}$) of 50 predicted images with their corresponding ground truths, respectively. The clustering of centroids within a specific scale range across all results indicates that the reconstructed images accurately represent the spatial location of the fast-rotating star relative to the ground truth.

\begin{figure*}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}\hfil
  \caption{The second-order central moments provide information about the size and shape of stellar objects.  Shown here are all the second-order central moments for ground truth and predicted images generated by the trained GAN. From left to right these are $\mu_{11},\mu_{20},\mu_{02}$.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}\hfil
  \caption{Shown here are all the third-order central moments for
    ground truth and predicted images generated by the trained GAN.
    They represent the skewness of the brightness distributions.  The
    panels in reading order show
    $\mu_{30},\mu_{03},\mu_{21},\mu_{12}$.}
  \label{fig:moments}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of fast-rotating stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

@@

\subsection{The reconstructed Parameters for object}
The centroids \((x_c, y_c)\) indicate only the center of the fast-rotating star and its spatial location in the image. In contrast, the second-order central moments determine the orientation, semi-major axis, and eccentricity relative to the source's center \citep{teague1980image}. These moment-based parameters fully describe the two-dimensional ellipse that fits the image data.

>>>>

\section{Images Reconstructed by the GAN}
In this section, we begin by discussing phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis in which multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of \(2 \cdot 10^{-4}\), a kernel size of 5×5, and equal noise percentages (alpha/beta = 1) applied to both the original and generated images. A batch size of 1 is used, and equal training is provided to both the Discriminator and the Generator.

\subsection{Predicted Image from the Trained GAN}
Fig.~\ref{fig:GAN} demonstrates the success of the GAN in training a model to reconstruct the images of  fast-rotating stars using their Intensity Interferometry (II) observation . The GAN was trained on the training datasets for 60,000 steps and subsequently tested on various validation datasets to produce predicted images of the stars. The training was performed on a CPU using two nodes, each with 96 threads, and required approximately nine hours. In Fig.~\ref{fig:GAN}, four combined images illustrate the GAN's performance in reconstructing the stars' shape, size, and brightness distribution using II.
\begin{itemize}
\item{The left panel shows the signals collected from six baselines, which serve as the input for the Generator during training.}
\item{The first middle panel displays the real image, or ground truth, which the Disciminator loss function uses to distinguish from the images generated by the Generator. During training, the GAN aims to keep minimum difference between these ground truth image and generated images.}
\item{The second middle panel presents the reconstructed, or predicted, image produced by the trained GAN, highlighting its success in image reconstruction.}
\item{The right panel shows the difference between the ground truth and the predicted image in the interferometric plane. The uniform background color of the frames suggests that the GAN model picks up white noise in the interferometric plane.}
\end{itemize}
The predicted images in Fig.~\ref{fig:GAN} yield encouraging results, accurately conveying visual information about the source's size, shape, and brightness distribution across its surface using only six baselines. However, further improvements can be achieved by increasing the number of telescopes to maximize coverage of the $(u, v)$ plane, making the existing and upcoming CTAO an ideal candidate for this approach.

@@

The raw moment $M_{ij}$ of an image is defined as \citep{hu1962visual}
\begin{equation}
	M_{ij} = \sum_{x} \sum_{y} x^i y^j I(x, y)
	\label{eqn:Mom}
\end{equation}
where $I(x,y)$ represents the intensity at pixel $(x,y)$. The zeroth order raw moment, or monopole, represents the total intensity of an image. It is computed by summing all pixel values across the image, yielding an overall intensity measure. In this context, analyzing the monopole provides the total flux of fast-rotating stars. According to Eq.~\eqref{eqn:Mom}, the monopole of an image is calculated as:
\begin{equation}
	M_{00} = \sum_{x} \sum_{y} I(x, y).
\end{equation}
The left figure of Fig.~\ref{fig:cen} displays the monopole values for 50 reconstructed images. The plot reveals a linear relationship between the monopole of the ground truth (real image) on the $x$-axis and that of the predicted (reconstructed) image on the $y$-axis, consistent across sources of varying shapes and sizes. This linearity confirms that the predicted images have an overall intensity (flux) that closely matches the ground truth. However, while the monopole effectively represents the total brightness, it does not provide information about the position, shape, size, or detailed brightness distribution of the fast-rotating stars. For these aspects, higher-order moments are necessary.

The center of mass of the sky image of any stellar object, is determined by its centroid, which provides the $x$ and $y$ coordinates representing the spatial position of the image. This centroid is computed using the first-order raw moments in conjunction with the monopole (the zeroth-order moment). The formulations for the centroid along the $x$ and $y$ directions are given by:
\begin{eqnarray}
&&x_c = \frac{M_{10}}{M_{00}} = \frac{\sum_{x,y} x \cdot I(x,y)}{\sum_{x,y} I(x,y)} \nonumber \\
&&y_c = \frac{M_{01}}{M_{00}} = \frac{\sum_{x,y} y \cdot I(x,y)}{\sum_{x,y} I(x,y)}
\end{eqnarray}
Here, \(I(x,y)\) represents the intensity at pixel \((x,y)\), \(M_{00}\) is the monopole (total intensity), and \(M_{10}\) and \(M_{01}\) are the first-order raw moments along the $x$ and $y$ axes respectively. This formulation accurately captures the spatial center of mass of the stellar object in the image.

The middle and right panel of Fig.~\ref{fig:cen} compare the centroids ($x_{c}$, $y_{c}$) of 50 predicted images with their corresponding ground truths respectively. The clustering of centroids within a specific scale range across all results indicates that the reconstructed images accurately represent the spatial location of the fast-rotating star relative to the ground truth.

\begin{figure*}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}\hfil
  \caption{The second-order central moments provide information about the size and shape of stellar objects.  Shown here are all the second-order central moments for ground truth and predicted images generated by the trained GAN. From left to right these are $\mu_{11},\mu_{20},\mu_{02}$.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}\hfil
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}\hfil
  \caption{Shown here are all the third-order central moments for
    ground truth and predicted images generated by the trained GAN.
    They represent the skewness of the brightness distributions.  The
    panels in reading order show
    $\mu_{30},\mu_{03},\mu_{21},\mu_{12}$.}
  \label{fig:moments}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of the stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

@@

\subsection{The reconstructed Parameters for object}
The centroids \((x_c, y_c)\) indicate only the center of the star and its spatial location in the image. In contrast, the second-order central moments determine the orientation, semi-major axis, and eccentricity relative to the source's center \citep{teague1980image}. These moment-based parameters fully describe the two-dimensional ellipse that fits the image data.
